Reanalysis Step 4: Propensity Score Analysis with Twang

### 2025-08-31 reanalysis-4-propensity_scores.qmd
deps

# Function to process a single imputation
process_imputation <- function(i) {
  # Check for cached result
  cache_file <- file.path(cache_dir, paste0("twang_imp_", i, ".rds"))

  if (use_cache && file.exists(cache_file)) {
    cached_result <- readRDS(cache_file)
    weights <- cached_result$weights

@@

    # Run twang ps model - following analysis.Rmd approach
    twang_model <- ps(
      formula = ps_formula,
      data = data.frame(imputed_datasets[[i]]),  # Convert to data.frame as in analysis.Rmd
      n.trees = n_trees,
      interaction.depth = 3,
      shrinkage = 0.01,
      estimand = "ATT",
      stop.method = c("es.mean"),  # Using es.mean only for speed
      verbose = FALSE
    )

@@

# Use mclapply on Unix/Mac, parLapply on Windows
if (.Platform$OS.type == "unix") {
  # Unix/Mac - use mclapply
  parallel_results <- mclapply(
    1:n_imputations,
    process_imputation,
    mc.cores = n_cores
  )
} else {
  # Windows - use parLapply
  cl <- makeCluster(n_cores)

  # Export necessary objects to cluster
  clusterExport(cl, c(
    "cache_dir", "use_cache", "ps_formula", "imputed_datasets",
    "n_trees", "ps", "get.weights"
  ))

@@

  parallel_results <- parLapply(cl, 1:n_imputations, process_imputation)

>>>>

# Function to process a single imputation
process_imputation <- function(i, imp_datasets, formula, trees, cache_path, use_caching) {
  # Recreate variables inside function to avoid scoping issues
  library(twang)

  # Check for cached result
  cache_file <- file.path(cache_path, paste0("twang_imp_", i, ".rds"))

  if (use_caching && file.exists(cache_file)) {
    cached_result <- readRDS(cache_file)
    weights <- cached_result$weights

@@

    # Get the data for this imputation
    imp_data <- as.data.frame(imp_datasets[[i]])

    # Ensure lapse is properly formatted as numeric 0/1
    if ("lapse" %in% names(imp_data)) {
      if (is.factor(imp_data$lapse)) {
        # Convert factor to numeric via character to avoid level issues
        imp_data$lapse <- as.numeric(as.character(imp_data$lapse))
      } else if (is.character(imp_data$lapse)) {
        imp_data$lapse <- as.numeric(imp_data$lapse)
      } else if (is.logical(imp_data$lapse)) {
        imp_data$lapse <- as.numeric(imp_data$lapse)
      }

      # Ensure values are 0 or 1
      if (any(!imp_data$lapse %in% c(0, 1, NA))) {
        stop(paste("Lapse variable contains values other than 0, 1, or NA in imputation", i))
      }
    } else {
      stop(paste("Lapse variable not found in imputation", i))
    }

    # Run twang ps model
    twang_model <- ps(
      formula = formula,
      data = imp_data,
      n.trees = trees,
      interaction.depth = 3,
      shrinkage = 0.01,
      estimand = "ATT",
      stop.method = c("es.mean"),
      verbose = FALSE
    )

@@

# Use mclapply on Unix/Mac, parLapply on Windows
if (.Platform$OS.type == "unix") {
  # Unix/Mac - use mclapply
  parallel_results <- mclapply(
    1:n_imputations,
    function(i) process_imputation(i, imputed_datasets, ps_formula, n_trees, cache_dir, use_cache),
    mc.cores = n_cores
  )
} else {
  # Windows - use parLapply
  cl <- makeCluster(n_cores)

  # Export necessary objects to cluster
  clusterExport(cl, c(
    "process_imputation", "imputed_datasets", "ps_formula", 
    "n_trees", "cache_dir", "use_cache"
  ), envir = environment())

@@

  # Create wrapper function for parLapply
  parallel_results <- parLapply(cl, 1:n_imputations, function(i) {
    process_imputation(i, imputed_datasets, ps_formula, n_trees, cache_dir, use_cache)
  })
